```{r, include = FALSE}

knitr::opts_chunk$set(eval=FALSE)

```


# UI

## ui.R

Here we define the user interface (UI) for the dashboard. First we source the ui code for each of the tabs: 

```{r}

# source tabItems
source("TabItems/Tab_KPI.R")
source("TabItems/Tab_Taxa_Explorer.R")
source("TabItems/Tab_Student_Engagement.R")
source("TabItems/Tab_About.R")
source("TabItems/Tab_Record_Finder.R")


```

1. **Header Section (`dashboardHeader`):**
   - Sets up the header of the dashboard with the title "Biodiversity Monitoring Programme."

```{r}
# 1. HEADER ---------------------------------------------------------------------------------------------------

header <- dashboardHeader(
  title = "Biodiversity Monitoring Programme",
  titleWidth = 320
)

```


2. **Sidebar Section (`dashboardSidebar`):**
   - Creates a sidebar that contains a menu (`sidebarMenu`) with several items.
   - It includes a dropdown (`selectInput`) for choosing the survey year.
   - Menu items include "About," "Key Performance Indices," "Taxa Explorer," "Student Engagement," and "Record Finder."
   - Conditional panels (`conditionalPanel`) show/hide additional input fields based on the selected menu item. For example, when "Taxa Explorer" is selected, a dropdown to select a taxa is displayed; when "Student Engagement" is selected, a dropdown to select a student is displayed.


```{r}

# 2. SIDEBAR ---------------------------------------------------------------------------------------------------

sidebar <- dashboardSidebar(
  sidebarMenu(
    id = "sidebar",
    
    # select survey year
    selectInput(inputId = "year", label = "Select year", choices = years),
    
    # About Page
    menuItem("About", tabName = "about", icon = icon("circle-info")),
    
    # first menu item: Key Performance Indices
    menuItem("Key Performance Indices", tabName = "kpi", icon = icon("gauge")),
    
    # second menu item: Taxa Explorer
    menuItem("Taxa Explorer", tabName = "taxa_expl", icon = icon("crow")),
    
    # Show panel only when taxa explorer sidebar is selected
    useShinyjs(),
    
    # select taxa
    div(
      id = "taxa_cond",
      conditionalPanel(
        "input.sidebar == 'taxa_expl'",
        selectInput(inputId = "taxa_select", label = "Select taxa", choices = NULL)
      )
    ),
    
    # third menu item: Student Engagement
    menuItem("Student Engagement", tabName = "stud_expl", icon = icon("graduation-cap")),
    
    # select Observer
    div(
      id = "student_cond",
      conditionalPanel(
        "input.sidebar == 'stud_expl'",
        selectInput(inputId = "student_select", label = "Select Student", choices = NULL)
      )
    ),
    
    # fourth menu item: Record & Photos
    menuItem("Record Finder", tabName = "record_finder", icon = icon("camera-retro"))
  )
)

```

3. **Body Section (`dashboardBody`):**
   - Defines the main body of the dashboard.
   - Utilizes the `dashboardBody` function and includes additional styling using `tags$head` to link an external stylesheet (`custom.css`).
   - Uses the `tabItems` function to organize the content into different tabs (e.g., `Tab0`, `Tab1`, etc.).

```{r}

# 3. BODY ---------------------------------------------------------------------------------------------------

body <- dashboardBody(
  useShinyjs(),
  tags$head(
    tags$link(rel = "stylesheet", type = "text/css", href = "custom.css")
  ),
  tabItems(
    Tab0,
    Tab1,
    Tab2,
    Tab3,
    Tab4
  )
)


```


4. **UI Function (`dashboardPage`):**
   - Combines the header, sidebar, and body into a complete UI using the `dashboardPage` function.
   - This UI function essentially wraps up all the UI components, creating the structure of the Shiny app.

```{r}

# 4. UI Function ---------------------------------------------------------------------------------------------------

ui <- dashboardPage(header, sidebar, body)

```

## Tabs 

### Tab_About
### Tab_KPI
### Tab_Record_Finder
### Tab_Student_Engagement








